---
title: "Spotify Music Exploratory Data Analysis"
subtitle: "ETC5521 Assignment 2"
team: wallaby
author: 
  - Helen Evangelina
  - Rahul Bharadwaj
date: "`r Sys.Date()`"
bibliography: references.bib
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(devtools)
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(ggplot2)
library(skimr)
library(knitr)
library(kableExtra)
library(forcats)
library(broom)
library(gridExtra)
library(genius)
library(tidytext)
library(textdata)
library(DT)
library(corrplot)
library(ggthemes)
#to install the spotifyr package
#install.packages("devtools")
#devtools::install_github('charlie86/spotifyr')
library(spotifyr)
library(ggridges)
library(visdat)
```

```{r read-data, eval=FALSE}
readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv') %>% 
  write.csv("spotify_songs.csv")
```

```{r read-csv}
spotify_songs <- read.csv(here::here("spotify_songs.csv"))
spotify_songs <- spotify_songs %>% 
  mutate(track_artist = replace(track_artist, track_artist == "Ti<U+00EB>sto", "Tijs Michiel Verwest OON"))
```

# Music characteristics overtime

(put here or on the introduction?)

* It has been previously discussed about the different music components and the correlation between each, and also `track_popularity`. Now, another thing that we can look at is the trend of the music components overtime. Along time, more musical instruments and more genres are being introduced, changing our music taste. Therefore, analysing the trend of music components over time is an interesting thing to look at as it would be beneficial to understand how the characteristics of music are changing. As music evolves, we want to look at how the characteristics evolve. Are the music characteristics in 1957 similar to those in 2019? The trend of music genres has been analysed before, which shows that the trend is changing. Therefore, looking at the trend of music components would broaden the analysis more to see if the components are also changing.

```{r}
characteristics_mean <- spotify_songs %>%
 select(danceability,
        acousticness,
         duration_ms,
        energy,
        instrumentalness,
        key,
        liveness,
        loudness,
        mode,
        speechiness,
        tempo,
        valence,
        year) %>%
  pivot_longer(-year, names_to = "components",
               values_to = "values") %>%
  group_by(year,
           components) %>%
  summarise(mean = mean(values)) %>%
    group_by(year)

characteristics_mean$year <- as.Date(characteristics_mean$year, format = "%Y")
```


```{r components-trend, fig.cap="The trend of music components over the years."}
char_long <- spotify_songs %>%
  select(danceability,
        acousticness,
         duration_ms,
        energy,
        instrumentalness,
        key,
        liveness,
        loudness,
        mode,
        speechiness,
        tempo,
        valence,
        year) %>%
  pivot_longer(-year, names_to = "components",
               values_to = "values")

 ggplot() +
    geom_point(data = char_long,
               aes(x = as.Date(year, format = "%Y"),
                 y = values,
                 color = components,
                 alpha = 0.5)) +
   
   geom_line(data = characteristics_mean,
             aes(x = year, 
                y = mean),
             size = 0.8) +
   geom_smooth(data = characteristics_mean,
             aes(x = year, 
                y = mean),
                color = "red") +
  facet_wrap(~components, scales = "free") +
   theme(legend.position = "none") +
   ggtitle("Trend of Music Components Overtime") +
  xlab("Year") +
  ylab("Values")
```


To answer the research question "How is the music characteristic over time and how is it changing?", we plot the values of the different music components against year, and faceted according to the components. We can see from Figure \@ref(fig:components-trend) that there are some components which are changing over time.

Acousticness, mode, tempo and valence tend to decrease over time, while danceability and energy tend to increase. The decreasing valence indicates that new songs tend to be sadder as more and more unhappy songs are being released every year. "Happiness and brightness in music has declined, while sadness increased in the last 30 years or so" (https://entertainment.inquirer.net/274757/people-prefer-happy-music-sad-songs-trend-past-30-years-study). Researches have suggested that the usage of positive emotions has declined. However, there is a high variability in valence values which means that there are varying types of songs. The increasing danceability and energy is due to the rising of electronic music. 

Other music components tend to follow a steady trend overtime. However, looking at instrumentalness, speechiness, liveness, and loudness, there has been more variances along time. Loudness tend to range between -20 to 0 decibels, except for some outliers present where the loudness is really low.

 
The emergence of more modern music with high energy and


The music characteristics might change because of the emergence of new types and more modern music. But is the music characteristic also changing for an artist? It would be interesting to look at the trend to see if one specific artist also has a shift in his music characteristics. Is the same artist making the same kind of music through time?

To answer this question, we are going to look at the top five artists and look at their music characteristics trend over time. We are not using the `track_popularity` to determine the top five artists here as some artists that are in the top five only have songs in a certain year, thus we are not able to compare the music characteristics over time. Therefore, we are using number of `track_name` instead. The following table display the top five aartists:

```{r topartists}
topartists <- spotify_songs %>% 
  group_by(track_artist) %>%
  summarize(Total = n_distinct(track_name)) %>% 
  arrange(desc(Total))  

head(topartists, 5) %>%
  kable()
```
Unlike the other four artists, Queen has a different music timeline. Therefore, for the purpose of having a clearer visualisation, we split Queen with the others.


```{r top4-wrangling}
top4artists <- spotify_songs %>%
  filter(track_artist == c("Martin Garrix", "David Guetta", "Logic", "Hardwell")) %>%
  select(danceability,
        acousticness,
        energy,
        instrumentalness,
        key,
        liveness,
        loudness,
        mode,
        speechiness,
        tempo,
        duration_ms,
        valence,
        year,
        track_artist)

top4_components <- top4artists %>%
  pivot_longer(danceability:valence, names_to = "components",
               values_to = "values") %>%
  group_by(year,
           track_artist,
           components) %>%
  summarise(mean = mean(values))

top4_components$year <- as.Date(top4_components$year, format = "%Y")         
```

```{r top4-components, fig.cap="The musical components of the top 5 artists (excluding Queen) over time."}
top4_components %>%
  group_by(year, track_artist) %>%
  ggplot() +
  geom_line(aes(x = year, 
                y = mean,
                color = track_artist)) +
  facet_wrap(~components, scales = "free") +
  ggtitle("Trend of music components of the top artists") +
  ylab("Values") +
  xlab("Year")
ggplotly()
```

It can be seen from Figure \@ref(fig:top4-components) that music components of the individual artist shift over time. Some components are pretty volatile. David Guetta, who (berkarir) in both pop and edm, has a huge change in his danceability and energy. He tends to produce more music with lower danceability but higher energy these days. His songs also have lower duration compared to those in the past. The valence of his songs change overtime too - with 2010 and 2011 being the years when he produced more positive songs, and his songs tend to get less positive over the years. Hardwell focuses on edm, which explains the low speechiness. His instrumentalness used to be pretty high in 2013, however it is shifting towards zero over time. Unlike David Guetta, Hardwell tends to produce more positive songs these days.

Being a rapper, it makes sense that Logic has a very close to zero instrumentalness. The trend of key component in his music follows a decreasing trend over time. The valence, acousticness, and liveness of his songs are pretty volatile over time. One inetersting thing about Hardwell is that his danceability and tempo were decreasing in around 2016, indicating that he produced less upbeat and slower tempo music in this year. After 2016, his danceability and tempo are increasing again. Martin Garrix is an edm artist. He has a decreasing energy, loudness and duration over time. In 2013 and 2014, he produced music with a relatively high instrumentalness, however it dropped down across the years. Similar to Hardwell, Martin Garric creates more positive songs along the years - indicated by the increasing valence.

A clear thing to notice here is that all of them have a relatively low speechiness as they are all in either edm or rap genre. The most volatile component of all is mode.

```{r queen-wrangling}
queen <- spotify_songs %>%
  filter(track_artist == "Queen") %>%
  select(danceability,
        acousticness,
        energy,
        instrumentalness,
        key,
        liveness,
        loudness,
        mode,
        speechiness,
        tempo,
        duration_ms,
        valence,
        year,
        track_artist)

queen_components <- queen %>%
  pivot_longer(danceability:valence, names_to = "components",
               values_to = "values") %>%
  group_by(year,
           components) %>%
  summarise(mean = mean(values))

queen_components$year <- as.Date(queen_components$year, format = "%Y")         
```

```{r queen-trend, fig.cap="The musical components of Queen over time."}
queen_components %>%
  group_by(year) %>%
  ggplot() +
  geom_line(aes(x = year, 
                y = mean),
            color = "red") +
  facet_wrap(~components, scales = "free") +
  ggtitle("Trend of music components of Queen") +
  ylab("Values") +
  xlab("Year")
```

Figure \@ref(fig:queen-trend) is showing the music components of Queen's songs over time. And here we can see that the music characteristics evolve overtime. Queen is a rock band, hence explains the relatively high energy level. A very interesting thing here is that the danceability, energy, liveness, loudness, and speechiness are all dropping in 1992, while the acousticness is increasing sharply compared to the years before. This is due to Queen only produced one song in 1992 - the popular "We Are The Champions". This song is pretty different compared to other songs Queen produced as this song has less energy and less danceable. 

It can be concluded that the music characteristics evolve over time. And despite the changes, each artist has their own uniqueness in terms of the music they produce.




i wanna look at the relo between track artist and their musical characteristics(x as artist, y as characteristics?)

can also look at relationship between year (decade) and characteristics and make scatter plot matrix coloured by artists.

a <- top5artists %>%
  mutate(decade = round(as.numeric(year) - 4.5, -1)) %>%
   pivot_longer(danceability:valence, names_to = "characteristics",
               values_to = "values")
a$year <- as.Date(a$year, format = "%Y")

aa <- characteristics_topartists %>%
  group_by(year, track_artist)

ggplot() +
  geom_line(data = aa,
              aes(x = year, 
                y = mean,
                color = track_artist)) +
  geom_point(data = a,
             aes(x = year,
                 y = values,
             color = track_artist)) +
  facet_wrap(~characteristics, scales = "free")

the interesting one here is valence - so we look at it closer.


TO see if the characteristics are changing over time, can look at the correlation between the char and year.



```{r}
lm_year <- lm(year ~ acousticness, data = spotify_songs)

lm_dance <- lm(year ~ danceability, data = spotify_songs)

result <- as.data.frame(broom::tidy(lm_year)) %>% 
  filter(p.value < 0.05)

kable(result, digits = 2, caption = "lm(year~features)", booktabs = T, align = "c", escape = F) %>% 
  kable_styling(latex_options = c("striped", "hold_position"), full_width = F, position = "center") 

lm_year
lm_dance
```


# Conclusion

After Exploratory Data Analysis, our group got the answers to those questions. 
First of all, there is a positive or negative correlation between audio features and track popularity. However, as we all know, the value of a art work can't be measured only by numbers. The popularity of music artworks depends more on the artist's own popularity, creative talent or singing ability, or external factors such as world trends. The probability of success by deliberately catering to audio features and creating specific songs is not sufficient. 

Secondly, each top artist has its own artistic characteristics, and will be loved by specific groups of people. Top artists do not create music artworks according to the trend, instead, they will create their own trend for the world. 

As for the six kinds of music genres that can stand out from the modern music, there are also their own characteristics inside. It's hard to understand the reasons for their success because of their unique styles. What we can do is to determine the genre of each song according to its style. 

Finally, Although Coldplay, as one of the representative rock artist, their works contain more negative emotions. This is also in line with the rebellious and critical spirit of rock music, and this spirit has been respected by young people of different races all the time. They stick to their own style, try unconventional music routines as far as possible, and point to people's hearts with straightforward, profound and moving melody. This also confirms our analysis that Coldplay songs' lyrics convey negative emotions, which does not affect their popularity, but makes them top artists. In conclusion, track popularity will pay more attention to the singer's own ability and attitude, rather than audio features. The biggest role of audio features is to reflect the singer's music style, rather than increase popularity.

The R packages we used in this report: 
_@ggplot2, @skimr, @corrplot2017, @ggthemes, @DT, @devtools, @readr, @tidyverse, @dplyr, @lubridate, @knitr, @kableExtra, @broom, @gridExtra, @genius, @tidytext, @textdata, @spotifyr,  @ggridges ._

# Reference

