---
title: "ETC5521 Assignment 1"
subtitle: "Spotify Music Exploratory Data Analysis"
team: wallaby
author:
  - Chengzhi Ye
  - Yan Ma
date: "`r Sys.Date()`"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(ggplot2)
library(spotifyr)
library(skimr)
```


[This assignment is for ETC5521 Assignment 1 by Team `r rmarkdown::metadata$team` comprising of `r knitr::combine_words(rmarkdown::metadata$author)`.]{style="color:#006DAE;"}


# Introduction and motivation

Music, in a broad sense, is any art composed of sound, but it can express people's thoughts and thoughts, which implies the author's life experience, thoughts and feelings, and can bring people the enjoyment of beauty and the expression of human feelings. At the same time, music is also a form of social behavior, through which people can exchange feelings and life experiences. 

In ancient times, when the court held a banquet, or some talented people visited the landscape, they would play music to boost the fun. But in modern times, because the threshold of classical music is too high, and its development has gradually reached the extreme, it has become a very small group, while pop music (the general name of popular songs, including Rock, R&B, Latin, etc)  is gradually showing its own characteristics. Therefore, modern songs are quietly occupying the top position in people's hearts because of their outstanding performance in conveying emotion and life experience. Listening to pop music has also become the most common behavior in everyone's daily entertainment.

Nowadays, music plays an important role in people's life. It plays an indispensable role in helping people manage and improve their quality of life. As fans of music, we not only enjoy music, but also wonder how music strikes people's hearts with simple tones, rhythms, timbres and words. How high is the position of genre in music performance? How much influence does the genre, or the various attributes of songs, have on music? Where do we like music? Whether it makes us dance or sing unconsciously, or does it convey our emotions and implicate our thoughts? All these are the motivations that we continue to study. But now listening software has sprung up like mushrooms. After careful consideration, our group decided to select Spotify as the research object. First of all, let me introduce Spotify.

Spotify is a legitimate streaming music service platform, which has been supported by Warner Music, Sony, EMI and other major record companies around the world. Now it has more than 60 million users, and it is the world's leading large-scale online streaming music playing platform.

Because Spotify contains a large number of users' data, four users who are very interested in it, Charlie Thompson, Josiah parry, Donal Phipps, and Tom Wolff decided to make it easier for everyone to know their own preferences or the mainstream of most people's listening to songs through spotify's API, thus creating Spotifyr package. Also, it is the source of our group assignment data.

In addition to Spotify package, our data is also mixed with blog post data created by Kaylin Pavlik. Six main categories (EDM, Latin, pop, R&B, rap, rock) are used to classify 5000 songs. The combination of the two data has a great effect on the study of the popularity of pop music.

# Data description
```{r read-data, eval=FALSE}
spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv') %>% 
  write.csv("spotify_songs.csv")
```

```{r read-csv}
spotify_song <- read.csv(here::here("spotify_songs.csv"))
```
         

## Data Source

The data of this report is part of the [tidytuesday chanllenge](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-21/readme.md), which comes from [Spotify](https://www.spotify.com/) via the  [spotifyr](https://www.rdocumentation.org/packages/spotifyr/versions/2.1.1) package.


## Data Structure

The **variables** in this dataset are `r names(spotify_songs)`, time frame of collection is from 1957-01-01 to 2020-01-29.

**Data collection methods:** Spotifyr package can extract track audio characteristics or other related information from Spotify's Web API in batches. For example, if you want to search for an artist, just type in his name, and all his albums or songs will be listed in seconds. Meanwhile, Spotifyr package will record the popularity metrics of all tracks or albums, so it is easy to understand the correlation between music popularity and music characteristics. Then, Jon Harmon and Neal Grantham extracted the Spotifr package and added the content of Kaylin Pavlik's recent blogpost to divide the genre of nearly 5000 songs, thus generating the Tidytuesdayr package we need for this assignment.

We chose music works created by artists that can be found on Spotify from January 1, 1957 to January 29, 2020.

Here is a brief summary of the **data structure**:

```{r data-structure}
skimr::skim(spotify_song) 
```
The spotify_song is tabular data, which contains 24 columns and 32,833 rows. The variables of the dataset include "track_id", "track_name", "track_artist", "track_popularity", "track_album_id", "track_album_name", "track_album_release_date", "playlist_name", "playlist_id", "playlist_genre"  "playlist_subgenre", "danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "tempo" and "duration_ms".     


# Analysis and findings

[FILL] Should include at least one plot or numerical summary for each of your questions, that helps the reader arrive at an answer. You should also write paragraphs describing the methods, summaries and findings. 

## Potential analysis questions

Research Question:

Qusetion 2, 3, 4:

1. Characteristics of each type of song: (col 13:24)
  Summarise by type (genre) of music: (playlist_name, playlist_genre, playlist_subgenre)

2. What type of songs are popular?
  y ~ x(features)

3. Each type of song was populated in what time period? (track_album_release_date)

4. Analyse my favorate artist, eg: Maroon 5.


## Limitation

1. popularity: too many zeros & how to calculate it


```{r}
spotify_song %>% 
  filter(track_artist == "Maroon 5")
```


# References




